Modular flow and holographic entanglement entropy in cutoff AdS / \TTbar deformed theory

### 2021-06-22 alpha/CutoffModularFlow/notes.tex
Update the boundary story in notes

\usepackage{xspace}
\newcommand{\TTbar}{\ensuremath{T\bar{T}}\xspace}

@@

\section{Entanglement in the boundary story}
	A boundary interval $\mcal{A}$ defines a denisty matrix:
	\begin{equation}
		\rho_\mcal{A}
		= \tr_{\bar{\mcal{A}}} \rho
		= e^{-\mcal{H}_\mcal{A}}
	\end{equation}
	where $\bar{\mcal{A}}$ is the complement of $\mcal{A}$ and $\mcal{H}_\mcal{A}$ is the corresponding modular Hamiltonian. $\mcal{H}_\mcal{A}$ is generically nonlocal, but in a highly symmetric background (e.g.~AdS vacuum) it might be geometrically realized and corresponds to a symmetry of the underlying quantum field theory. This is the modular flow $\xi$.

	In such case there is a generalized Rindler transformation $f$ that maps the casual domain $\mcal{D}$ to a patch $\{(\tau,\tilde{x})\}$ in some noncompact generalized Rindler spacetime, with $\pdd{\tau}$ an isometry of the spacetime. Then we have:
	\begin{equation}
		\xi = 2\pi \pdd{\tau}
		= 2\pi \sum_i s^i H_i,
	\quad
		\tau \sim \tau + 2\pi i
		= \tau + i\beta_\tau,
	\quad
		\beta_\tau = 2\pi
	\end{equation}
	Here $H_i$ are the symmetry generators. The causal domain of dependence $\mcal{D}$ is mapped to some thermal state with normalized inverse temperature $\beta_\tau = 2\pi$. Here we choose to absorb such $2\pi$ into the definition of $\xi$, so the Euclidean periodicity w.r.t.~$\xi$ is simply normalized to 1, which means the surface gravity along the horizon is given by:
	\begin{equation}
		\kappa_\xi = 2\pi
	\end{equation}
\subsection{Boundary modular flows}
	In usual $\mrm{CFT}_2$, this prescription gives us the modular flow generators for a single interval $\mcal{A}$:
	\begin{equation}
		\xi_\mrm{CFT} = 2\pi \sum_{n=0,\pm 1} \pqty{
				s_n L_n - \bar{s}_n \bar{L}_n
			},
	\end{equation}
	\begin{equation}
		s_{-1} = \frac{u_+ u_-}{u_+ - u_-},\quad
		s_0 = - \frac{u_+ + u_-}{u_+ - u_-},\quad
		s_{+1} = \frac{1}{u_+ - u_-}
	\end{equation}
	Where the $H_i$'s are given by $L_n,\bar{L}_n$ with $n=0,\pm 1$, i.e.~the global conformal generators, and $(u,v)_\pm$ or $(t,x)_\pm$ marks the endpoints of the interval $\mcal{A}$. The overall coefficient $2\pi$ is fixed by requiring that $\kappa_\xi = 2\pi$.

	For simplicity, we now restrict to an interval centered at the origin: $
		u_\pm v_\pm
		= x^2_\pm - t^2_\pm
		= (\frac{{\tilde L}}{2})^2
	$, where $x_\pm = -x_\mp$, and same for $t_\pm, u_\pm, v_\pm$. 
	We then introduce the \textit{rapidity} $\theta$ to nicely parametrized the boosted interval. For a constant time slice, we have $\theta = 0$. For a general boosted interval, we have:
	\begin{equation}
		x_\pm = \pm \frac{{\tilde L}}{2} \cosh \theta,
	\quad
		t_\pm = \pm \frac{{\tilde L}}{2} \sinh \theta,
	\end{equation}
	Note that the RT surface lies on the plane $
		t_\pm / x_\pm = \tanh \theta
	$, namely $\theta$ is constant along the RT surface. 
	We have:
	\begin{gather}
	\begin{aligned}
		\xi_\mrm{CFT} &= \frac{2\pi}{{\tilde L}} \bigg\{
			\pqty\Big{
				\pqty{
					(\tfrac{{\tilde L}}{2})^2
					- (t^2 + x^2)
				} \cosh\theta
				+ 2tx \sinh\theta
			} \,\pdd{t}
		\\ &\qquad\qquad 
			+ \pqty\Big{
				\pqty{
					(\tfrac{{\tilde L}}{2})^2
					+ (t^2 + x^2)
				} \sinh\theta
				- 2tx \cosh\theta
			} \,\pdd{x}
		\bigg\}
	\end{aligned}
	\end{gather}
	Suppose the interval $\mcal{A}$ lies on the constant $t = 0$ slice; around the constant $t = 0$ slice, the modular flow becomes:
	\begin{equation}
		\xi_\mrm{CFT}|_{t = 0} 
		= \frac{2\pi}{{\tilde L}}
			\pqty{
				\pqty{\frac{{\tilde L}}{2}}^2
				- x^2
			} \,\pdd{t}
	\end{equation}

	However, after the \TTbar-deformation, the scaling and special conformal transformations are no longer symmetries of the theory, so we cannot simply take $\xi_\mrm{CFT}$ as the modular flow in the deformed theory.

	On the other hand, we still expect the basic properties of the modular flow to hold after the deformation; for example, we expect the endpoints $\pd\mcal{A}$ of $\mcal{A}$ to be fixed under the modular flow, and we suppose the $\xi_{T\bar{T}}$ should reduce back to $\xi_\mrm{CFT}$ as the deformation parameter $\lambda \to 0$. This leads us to the following proposal of $\xi_{T\bar{T}}|_{t = 0}$, for an interval $\mcal{A}$ with coordinate length $L$:

	Note that:
	\begin{enumerate}[topsep=0pt]
	\item The dependence of $\xi_{T\bar{T}}|_{t = 0}$ w.r.t.~$L$ is very similar to the dependence of $\xi_\mrm{CFT}|_{t = 0}$ w.r.t.~${\tilde L}$, except that now the normalization is slightly modified to include the deformation parameter $\lambda$. 
	\item This is not yet a proposal for the full modular flow, but only its values around the $t = 0$ slice. But we shall see that this is enough to give us the entanglement entropy of a \TTbar-deformed theory. 
	\end{enumerate}

@@

\subsection{Rindler map in finite temperature}
	One can use the method in (1909.13808) to calculate the entanglement entropy in the thermal case and finite size case from the field theory side. The Rindler transformation from the domain of dependence of the interval $(-l_p,l_p)$ to the sphere is
	\begin{equation}
	t=l_p\frac{\sin\theta \sinh(\tau/l_p)}{1+\sin\theta \cosh(\tau/l_p)},\ \  x=l_p\frac{\cos\theta}{1+\sin\theta\cosh(\tau/l_p)}
	\end{equation}
	the conformal factor reads,
	\begin{equation}
	\Omega=\frac{1}{1+\sin\theta\cosh(\tau/l_p)}
	\end{equation}
	At endpoints, it becomes $1$. From the discussion in (1909.13808), the contribution to the entanglement entropy totally comes from the endpoints of the interval, so one can use the such conformal transformation where the conformal factor at the endpoints is $1$ to calculate the entanglement entropy. Also, in the conformal map above from the casual diamond to the sphere, only the length of the interval matters. If these two endpoints lie at other position, one can use the rotation and translation symmetry to put it at $(0,\pm l_p)$, since the $T\bar T$ deformed theory also admits these symmetries. Now let's consider the thermal case, the conformal transformation from the thermal cylinder to the plane reads,
	\begin{equation}
	z=\frac{1}{r_1}e^{r_1w},\ \ \ r_1=\frac{2\pi}{\beta}
	\end{equation} 
	where $z$ is the coordinate on the plane, and $w$ is the coordinate on the thermal cylinder with identification,
	\begin{equation}
	w\sim w+i\beta
	\end{equation}
	The endpoints of the interval on the thermal cylinder is
	\begin{equation}
	w=\pm l
	\end{equation} 
	The length of the interval is $L=2l$. Under the conformal map, the endpoints on the plane is
	\begin{equation}
	\frac{1}{r_1}e^{\pm r_1 l}
	\end{equation}
	the length is $\frac{2}{r_1}\sinh(l r_1)$, so
	\begin{equation}
	l_p=\frac{1}{r_1}\sinh(l r_1)
	\end{equation}
	The conformal factor is
	\begin{equation}
	e^{r_1(w+\bar{w})}
	\end{equation}
	Note that compared to the usual map from cylinder to plane $z=e^{r_1 w}$, we put the overall scaling factor $\frac{1}{r_1}$ to make sure that the conformal factor at the endpoints is $1$. From 1806.07444, the entanglement entropy for $l_p$ is
	\begin{equation}
	S_p(l)=\frac{c}{3}ArcSinh\frac{l_p}{a}
	\end{equation}
	where
	\begin{equation}
	a^2=\frac{\mu c}{12}
	\end{equation}
	The C-function reads,
	\begin{equation}
	C(L)=L S'(L=2l)=\frac{c}{3}\frac{1}{\sqrt{1+4a^2/L^2}}
	\end{equation}
	Now in the thermal case, $l_p\rightarrow \frac{1}{r_1}\sinh(l r_1)$
	\begin{equation}
	S_t=\frac{c}{3}ArcSinh(\frac{1}{r_1 a}\sinh(l r_1))
	\end{equation}
	which is the same as the results obtained from the local temperature trick \eqref{} and the gravity calculation \eqref{}. Now consider the finite size case, the conformal transformation is,
	\begin{equation}
	z=\frac{1}{ir_1}e^{iwr_1},\quad
		r_1 = \frac{2\pi}{\lambda}
	\end{equation}
	where $w$ is the cylinder coordinate,
	\begin{equation}
	w\sim w+ \lambda
	\end{equation}
	The endpoints of the interval are 
	\begin{equation}
	w\pm l
	\end{equation}
	The conformal factor is
	\begin{equation}
	e^{ir_1(w+\bar{w})}
	\end{equation}
	which is $1$ at the endpoints. The plane length reads,
	\begin{equation}
	l_p=\frac{1}{r_1}\sin lr_1
	\end{equation}
	So the final result is
	\begin{equation}
	S_f=\frac{c}{3}ArcSinh(\frac{1}{r_1a}\sin(lr_1))
	\end{equation}

>>>>

\usepackage{xspace}
\newcommand{\TTbar}{\ensuremath{T\bar{T}}\xspace}
\newcommand{\arcsinh}{\ensuremath{\mop{arcsinh}}}

@@

\section{Entanglement in the boundary story}

\subsection{Rindler map in finite temperature}

\newcommand{\lp}{\ensuremath{l_p}}
\newcommand{\halfL}{\ensuremath{\frac{L}{2}}}

	One can use the method in \cite{Lewkowycz:2019xse} to calculate the entanglement entropy in the thermal case and finite size case from the field theory side. The Rindler transformation from the domain of dependence of the interval $(-\lp,\lp)$ to the sphere is
	\begin{equation}
	t=\lp\frac{\sin\theta \sinh(\tau/\lp)}{1+\sin\theta \cosh(\tau/\lp)},\ \  x=\lp\frac{\cos\theta}{1+\sin\theta\cosh(\tau/\lp)}
	\end{equation}
	the conformal factor reads,
	\begin{equation}
	\Omega=\frac{1}{1+\sin\theta\cosh(\tau/\lp)}
	\end{equation}
	At endpoints, it becomes $1$. From the discussion in \cite{Lewkowycz:2019xse}, the contribution to the entanglement entropy totally comes from the endpoints of the interval, so one can use the such conformal transformation where the conformal factor at the endpoints is $1$ to calculate the entanglement entropy. Also, in the conformal map above from the casual diamond to the sphere, only the length of the interval matters. If these two endpoints lie at other position, one can use the rotation and translation symmetry to put it at $(\pm \lp,0)$, since the $T\bar T$ deformed theory also admits these symmetries.

	Now let's consider the thermal case. The conformal transformation from the thermal cylinder to the plane reads,
	\begin{equation}
	z=\frac{1}{\kappa}e^{\kappa w},\ \ \ \kappa=\frac{2\pi}{\beta}
	\end{equation} 
	where $z$ is the coordinate on the plane, and $w$ is the coordinate on the thermal cylinder with identification,
	\begin{equation}
	w\sim w+i\beta
	\end{equation}
	The endpoints of the interval on the thermal cylinder is
	\begin{equation}
	w=\pm \halfL
	\end{equation} 
%	The length of the interval is $L=2l$. 
	Under the conformal map, the endpoints on the plane is
	\begin{equation}
	\frac{1}{\kappa}e^{\pm \kappa \frac{L}{2}}
	\end{equation}
	the length is $\frac{2}{\kappa}\sinh(\halfL \kappa)$, so one can use translations and rotations to map the endpoints to $(\pm \lp, 0)$ on the plane 
	\begin{equation}
	l_p=\frac{1}{\kappa}\sinh(\halfL \kappa)
	\end{equation}
	The conformal factor is
	\begin{equation}
	e^{\kappa(w+\bar{w})}
	\end{equation}
	Note that compared to the usual map from cylinder to plane $z=e^{\kappa w}$, we put the overall scaling factor $\frac{1}{\kappa}$ to make sure that the conformal factor at the endpoints is $1$.

	From \cite{Donnelly:2018bef}, the entanglement entropy for zero temperature is
	\begin{equation}
	S(L)=\frac{c}{3} \arcsinh \frac{\lp}{a}
	\end{equation}
	where
	\begin{equation}
	a^2=\frac{\mu c}{12}
	\end{equation}
	The $C$-function reads,
	\begin{equation}
	C(L)=L S'(L)=\frac{c}{3}\frac{1}{\sqrt{1+4a^2/L^2}}
	\end{equation}
	Now in the thermal case, $l_p\rightarrow \frac{1}{\kappa}\sinh(\halfL \kappa)$
	\begin{equation}
	S_t=\frac{c}{3}\arcsinh \pqty{
			\frac{1}{\kappa a}\sinh(\halfL \kappa)
		}
	\end{equation}
	which is the same as the results obtained from the local temperature trick \eqref{} and the gravity calculation \eqref{}. Now consider the finite size case, the conformal transformation is,
	\begin{equation}
	z=\frac{1}{ir_1}e^{iwr_1},\quad
		r_1 = \frac{2\pi}{\lambda}
	\end{equation}
	where $w$ is the cylinder coordinate,
	\begin{equation}
	w\sim w+ \lambda
	\end{equation}
	The endpoints of the interval are 
	\begin{equation}
	w\pm l
	\end{equation}
	The conformal factor is
	\begin{equation}
	e^{ir_1(w+\bar{w})}
	\end{equation}
	which is $1$ at the endpoints. The plane length reads,
	\begin{equation}
	l_p=\frac{1}{r_1}\sin lr_1
	\end{equation}
	So the final result is
	\begin{equation}
	S_f=\frac{c}{3}ArcSinh(\frac{1}{r_1a}\sin(lr_1))
	\end{equation}

\pagebreak
\subsection{Boundary modular flows}

	A boundary interval $\mcal{A}$ defines a denisty matrix:
	\begin{equation}
		\rho_\mcal{A}
		= \tr_{\bar{\mcal{A}}} \rho
		= e^{-\mcal{H}_\mcal{A}}
	\end{equation}
	where $\bar{\mcal{A}}$ is the complement of $\mcal{A}$ and $\mcal{H}_\mcal{A}$ is the corresponding modular Hamiltonian. $\mcal{H}_\mcal{A}$ is generically nonlocal, but in a highly symmetric background (e.g.~AdS vacuum) it might be geometrically realized and corresponds to a symmetry of the underlying quantum field theory. This is the modular flow $\xi$.

	In such case there is a generalized Rindler transformation $f$ that maps the casual domain $\mcal{D}$ to a patch $\{(\tau,\tilde{x})\}$ in some noncompact generalized Rindler spacetime, with $\pdd{\tau}$ an isometry of the spacetime. Then we have:
	\begin{equation}
		\xi = 2\pi \pdd{\tau}
		= 2\pi \sum_i s^i H_i,
	\quad
		\tau \sim \tau + 2\pi i
		= \tau + i\beta_\tau,
	\quad
		\beta_\tau = 2\pi
	\end{equation}
	Here $H_i$ are the symmetry generators. The causal domain of dependence $\mcal{D}$ is mapped to some thermal state with normalized inverse temperature $\beta_\tau = 2\pi$. Here we choose to absorb such $2\pi$ into the definition of $\xi$, so the Euclidean periodicity w.r.t.~$\xi$ is simply normalized to 1, which means the surface gravity along the horizon is given by:
	\begin{equation}
		\kappa_\xi = 2\pi
	\end{equation}

	In usual $\mrm{CFT}_2$, this prescription gives us the modular flow generators for a single interval $\mcal{A}$:
	\begin{equation}
		\xi_\mrm{CFT} = 2\pi \sum_{n=0,\pm 1} \pqty{
				s_n L_n - \bar{s}_n \bar{L}_n
			},
	\end{equation}
	\begin{equation}
		s_{-1} = \frac{u_+ u_-}{u_+ - u_-},\quad
		s_0 = - \frac{u_+ + u_-}{u_+ - u_-},\quad
		s_{+1} = \frac{1}{u_+ - u_-}
	\end{equation}
	Where the $H_i$'s are given by $L_n,\bar{L}_n$ with $n=0,\pm 1$, i.e.~the global conformal generators, and $(u,v)_\pm$ or $(t,x)_\pm$ marks the endpoints of the interval $\mcal{A}$. The overall coefficient $2\pi$ is fixed by requiring that $\kappa_\xi = 2\pi$.

	For simplicity, we now restrict to an interval centered at the origin: $
		u_\pm v_\pm
		= x^2_\pm - t^2_\pm
		= (\frac{{\tilde L}}{2})^2
	$, where $x_\pm = -x_\mp$, and same for $t_\pm, u_\pm, v_\pm$. 
	We then introduce the \textit{rapidity} $\theta$ to nicely parametrized the boosted interval. For a constant time slice, we have $\theta = 0$. For a general boosted interval, we have:
	\begin{equation}
		x_\pm = \pm \frac{{\tilde L}}{2} \cosh \theta,
	\quad
		t_\pm = \pm \frac{{\tilde L}}{2} \sinh \theta,
	\end{equation}
	Note that the RT surface lies on the plane $
		t_\pm / x_\pm = \tanh \theta
	$, namely $\theta$ is constant along the RT surface. 
	We have:
	\begin{gather}
	\begin{aligned}
		\xi_\mrm{CFT} &= \frac{2\pi}{{\tilde L}} \bigg\{
			\pqty\Big{
				\pqty{
					(\tfrac{{\tilde L}}{2})^2
					- (t^2 + x^2)
				} \cosh\theta
				+ 2tx \sinh\theta
			} \,\pdd{t}
		\\ &\qquad\qquad 
			+ \pqty\Big{
				\pqty{
					(\tfrac{{\tilde L}}{2})^2
					+ (t^2 + x^2)
				} \sinh\theta
				- 2tx \cosh\theta
			} \,\pdd{x}
		\bigg\}
	\end{aligned}
	\end{gather}
	Suppose the interval $\mcal{A}$ lies on the constant $t = 0$ slice; around the constant $t = 0$ slice, the modular flow becomes:
	\begin{equation}
		\xi_\mrm{CFT}|_{t = 0} 
		= \frac{2\pi}{{\tilde L}}
			\pqty{
				(\tfrac{{\tilde L}}{2})^2
				- x^2
			} \,\pdd{t}
	\end{equation}

	However, after the \TTbar-deformation, the scaling and special conformal transformations are no longer symmetries of the theory, so we cannot simply take $\xi_\mrm{CFT}$ as the modular flow in the deformed theory. 
	But we shall see that by properly specifying the integration range, this is enough to give us the entanglement entropy of a \TTbar-deformed theory.

@@
